Design and Implementation of Probabilistic Programming Language Anglican

### 2016-07-31 paper/paper.tex
example of factored continuation

\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{listings}
\usepackage[lighttt]{lmodern}
\ttfamily
\DeclareFontShape{OT1}{lmtt}{m}{it}
     {<->sub*lmtt/m/sl}{}
\lstdefinestyle{default}{frame=single,language=Anglican, basicstyle=\ttfamily\small, showstringspaces=false}

@@

\paragraph{Trampolining}

@@

To realize potential danger of code size explosion, consider
CPS transformation of code
\begin{lstlisting}[style=default]
(if (adult? person)
  (if (male? person)
     (choose-beer)
     (choose-wine))
  (choose-juice))
\end{lstlisting}
with continuation 
\begin{lstlisting}[style=default]
(fn [choice _]
  (case (kind choice)
    :beer (beer-jar choice)
    :wine (wine-glass choice)
    :juice (juice-bottle choice)))
\end{lstlisting}
The code of the continuation, represented by an \texttt{fn}
form, will be repeated three times. In general, CPS code can
grow extremely large if symbolic continuations are inserted
repeatedly.

To circumvent this inefficiency, CPS transformers for
expressions with multiple continuation points (\texttt{if} and
derivatives, \texttt{and}, \texttt{or}, and \texttt{case}) bind
the continuation to a fresh symbol if it is not yet a symbol.
Macro \texttt{defn-with-named-cont} establishes the binding
automatically.

>>>>

\usepackage{listings}
\usepackage[lighttt]{lmodern}
\ttfamily
\DeclareFontShape{OT1}{lmtt}{m}{it}
     {<->sub*lmtt/m/sl}{}
\lstdefinestyle{default}{language=Anglican, basicstyle=\ttfamily\small, showstringspaces=false}

@@

\paragraph{Managing stack size}

@@

\paragraph{Avoiding exponential code growth}

To realize potential danger of code size explosion, consider
CPS transformation of code
\begin{lstlisting}[style=default]
(if (adult? person)
  (if (male? person)
     (choose-beer)
     (choose-wine))
  (choose-juice))
\end{lstlisting}
with continuation 
\begin{lstlisting}[style=default]
(fn [choice _]
  (case (kind choice)
    :beer (beer-jar choice)
    :wine (wine-glass choice)
    :juice (juice-bottle choice)))
\end{lstlisting}
The code of the continuation, represented by the \texttt{fn}
form, will be repeated three times. In general, CPS code can
grow extremely large if symbolic continuations are inserted
repeatedly.

To circumvent this inefficiency, CPS transformers for
expressions with multiple continuation points (\texttt{if} and
derivatives, \texttt{and}, \texttt{or}, and \texttt{case}) bind
the continuation to a fresh symbol if it is not yet a symbol.
Macro \texttt{defn-with-{\linebreak[0]}named-{\linebreak[0]}cont}
establishes the binding automatically:
\begin{lstlisting}[style=default]
=> (cps-of-if '(c t f) '(fn [x] (* 2 x)))
(let [cont (fn [x] (* 2 x))]
  (if c
      (fn [] (cont t $state))
      (fn [] (cont f $state))))
\end{lstlisting}
